
While we ultimately decided to use the OverFeat software for classification,
we originally considered implementing our own algorithm as well.
The intention was that we could find something with better interpretability,
albeit with some loss of accuracy.
One approach we found for this, based on a paper by Coates and Ng,
was to use spherical k-means clustering for feature extraction.
In particular, each image in the training set is divided into 16 by 16 patches
of pixels, and spherical k-means clustering is applied to produce a dictionary of features.
To classify a test image, 
first each patch of the image is mapped to an entry in the feature dictionary.
Many different mappings are possible;
typically there is some tradeoff between efficiency and accuracy.
The number of occurences of each feature can then be counted---although this
is not the only approach---and provided to an apprporiately-trained linear
classifier.

This k-means-based approach to the problem had the potential to be easily
interpretable: our hope was that many patches in the feature dictionary would
have a clear correspondence to either cats or dogs.
Moreover, the original paper was flexible to the needs of the reader.
A wide array of possibilities for tuning and adjusting were presented,
which would have given us many avenues for improving accuracy.

Spherical k-means clustering is sufficiently exotic that we could find no
pre-existing libraries implementing it, so we wrote our own implementation.
A major concern was how well the algorithm would scale to our very large
training set, and we made preliminary investigations into moving our
Python code to Cython.
Due to the astounding success of OverFeat on the data set, and the fact that
many of our group members dropped out or were doing this for fun rather than
credit, this was as far as we got with implementing the k-means clustering
strategy.

Nevertheless, this project was a learning experience.
The spherical version of the k-means clustering algorithm was unfamiliar to
all of us, as were the damped k-means updates advocated in the paper.
Additionally, we learned that it's important to set clear project goals early
on; our group was somewhat directionless until halfway through the quarter.




